---
title: "ATI fDOM presentation"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    smooth_scroll: true 
    fig_width: 10
    code_folding: hide  
date: '2022-12-05'
---

# Overview

This document functions as a record of notes and code (i.e. scratch paper) for the Moorea Nutrient meeting in January 2023. Produced by Sean Swift following the various tasks thrown at him by Craig.

## Breakout 1

Groups 3 & 4:
- Update water column cluster analysis
- Silicate + Ocean Mixing proxy


### Discussion Notes

#### Nutrients affecting microbes & benthos

Nutrients don't actually affect heterotrophic microbes directly.
We would really really like benthic cover as a response...


#### fDOM qc 

Craig + Sean looked at matlab plots in jump side by side with R plots. Craig was concerned scrolling through plots from various Moorea projects. Looks like bright protein fluorescence and no DOM in samples e.g. Spiffy. Checked Spiffy sample 129 also for poor cropping of the Raman/Rayleigh scattering lines. 

**Conclusion**

- Same same, just R eem plots probably can't handle color scaling as well as JMP, which Craig is used to.
- When there's a bright spot in the protein/JP5 region, we can't see the swathes of fdom in the background.
- We can keep using R script outputs and trust them about the same as matlab
- Raman/Rayleigh crop issue is in both R and matlab


#### Organizing fDOM alongside other data

Nyssa produced a nice merged data.frame that is canonical across working groups.
Sean will join fDOM data onto this data.frame without dropping any rows.


```{r}
library(data.table)
library(ggplot2)

source("R/src/eem_plot_functions.R")
source("R/src/process_aqualog_functions.R")

```

#### Cross year comparison of fDOM

Craig: compare histograms of indices from 2021 and 2022


In JMP


Check two different fDOM parameters

Log CobleA
2021 = -1.4 to 2.4
2022 = -1.4 to 2.3

2022 contaminated = -.7 to -1.3

Contaminated = Hihg and bimodal distribution (other higher normal dist floating at higher values)


log CobleB
2021 = -1 to -2.5, median -1.4
2022 = -.6 to -2, median -1.45

2022 contaminated = -1 to -0.2


Note: these contaminated samples do not follow any 'natural' correlation with distance from shore, nutrient concentrations, etc. which also screams contamination

Until we figure out, probably have to exclude.


#### Check contam scans

Interpretation: all the DOM patterns in contaminated samples do look broadly similar

Also, we looked at Skye's notes for runs 6,7,8. We think all of these samples were shipped together in an ugly freezer box (see slack message Sean to Skye). Some of these were also urea capped scintillation vials.



```{r, fig.width=5, fig.height=4}
source("R/src/eem_plot_functions.R")
source("R/src/process_aqualog_functions.R")

# suspect samples based on Craig's clustering
# high in parameters across the board

exc = c(
"12_run6_1",
"121_run7_25",
"142_run4_11",
"147_run8_11",
"151_run8_15",
"152_run8_16",
"19_run6_4",
"205_run8_21",
"206_run8_22",
"27_run6_10",
"35_run6_16",
"44_run6_32",
"45_run6_33",
"58_run7_8",
"62_run7_12",
"63_run7_13",
"72_run7_19",
"99_run3_25",
"99_run3_25"
)

# plot contaminated samples in a single file

all_runs = readRDS("data/compiled/all_runs.rds")


# ati runs
ati_run_names = c(
    aroundtheisland_run1 = "ati",
    aroundtheisland_run2 = "ati",
    aroundtheisland_run3 = "ati",
    aroundtheisland_run4 = "ati",
    aroundtheisland_run5 = "ati",
    aroundtheisland_run6 = "ati",
    aroundtheisland_run7 = "ati",
    aroundtheisland_run8 = "ati"
)

ati_runs= all_runs[grepl("ati",names(all_runs))]


ati_eems_ll = sapply(1:length(ati_runs), function(i) ati_runs[[i]]$EEMs)
ati_eems = unlist(ati_eems_ll, recursive = F)

names(ati_eems) = gsub("_Group.+","", names(ati_eems))

# pull out contaminated eems
contam_eems = ati_eems[names(ati_eems) %in% exc]


contam_eem_ps = lapply(1:length(contam_eems), function(i) {
  an_eem = contam_eems[[i]]
  
  p = plot_eem(an_eem,
               sample_name = names(contam_eems)[i],
               rows_as_names = T)
    
  
  print(p)
  return(p)
  
})
  

```





#### Organize 2022 and 2021 fDOM data

'Join' (but actually do manually because of weirdness) the 2021 data sheet (pre-worked on by Tom), the 2022 data sheet (based on cap labels) and Nyssa's data sheet DataTurbWater.csv from google drive.

link: https://docs.google.com/spreadsheets/d/1ki_tV5Mm9EfV3jC-kTde5SC8gH11Zqc_KNZYioh5Nu0/edit?usp=sharing


1. Read in the data files

```{r}
# get clean data table from Nyssa
DataTurbWater = fread("data/clean/DataTurbWater.csv")

DTW_22 = DataTurbWater[Year == 2022]
DTW_21 = DataTurbWater[Year == 2021]



# get data from both years
mo22_fdom =  fread("data/clean/ati_fdom_clean.csv")

mo21_fdom = fread("data/clean/May2021_fdom_nutrients_turbinaria_compiled.csv")
```

2. Do some checking and merging


```{r}
# merge check function

check_site_names = function(fdom_sheet, clean_sheet, year){
  
  print(paste("Merge check for", year))
  print(paste("number of fdom samples = ", nrow(fdom_sheet)))
  print(paste("number of clean sheet samples = ", nrow(clean_sheet)))
  
# fdom sites in clean sheet
if(all(fdom_sheet$Site %in% clean_sheet$Site)){
 print(paste("all", year,"fdom sites are in the clean sheet :)")) 
}else{
  writeLines(c(paste(year, "fdom sites missing from clean sheet :("),
  fdom_sheet$Site[!(fdom_sheet$Site %in% clean_sheet$Site)]))
}

# clean sheet sites in fdom
if(all(clean_sheet$Site %in% fdom_sheet$Site)){
 print(paste("all clean sheet sites are in the", year," fdom sites :)"))
}else{
  writeLines(c(paste("clean sheet sites missing from", year,"fdom :("),
               clean_sheet$Site[!(clean_sheet$Site %in% fdom_sheet$Site)]))
}
}

```


```{r}
# Merging
# 2022 should merge well (names are pretty clean)
check_site_names(fdom_sheet = mo22_fdom, clean_sheet = DTW_22, year = "2022")

  
# 2021 may not merge well
check_site_names(fdom_sheet = mo21_fdom, clean_sheet = DTW_21, year = "2021")

```





```{r}
# fix the tables
## fix 22
fix_22_key = c(
  lter_3_bac = "LTER_3_Backreef",
  lter_3_fringe = "LTER_3_Fringe",
  lter_6_fringe = "LTER_6_Fringe",
  opu_1 = "OPU 1",
  opu_2 = "OPU 2"
)

fixed_mo22_fdom = copy(mo22_fdom)

fixed_mo22_fdom$Site[match(names(fix_22_key), fixed_mo22_fdom$Site)] = unname(fix_22_key)



check_site_names(fdom_sheet = fixed_mo22_fdom, clean_sheet = DTW_22, year = "2022")


# fix 21
fix_21_key = c(
  "Cook's 1" = "Cook 1",
  "LTER 3 Back Reef" =  "LTER_3_Backreef",
  "LTER 3 Fringe Reef" = "LTER_3_Fringe",
  "LTER 6 Fringe" = "LTER_6_Fringe"
)

fixed_mo21_fdom = copy(mo21_fdom)

fixed_mo21_fdom$Site[match(names(fix_21_key), fixed_mo21_fdom$Site)] = unname(fix_21_key)

# drop extra samples
extra_21fdom = c(
  1061,
  1062,
  1063
)

fixed_mo21_fdom = fixed_mo21_fdom[!(Site %in% extra_21fdom)]

# check again
check_site_names(fdom_sheet = fixed_mo21_fdom, clean_sheet = DTW_21, year = "2022")




```

Great! Site names are nice and clean.

Next up, let's pick our favorite fdom parameters. We're going to go with the naming conventions from the mo22 dataset (i.e. the index codes like Coble B instead of descriptors like 'marine humic like').


```{r}

# match index names between years 2021 and 2022
name_key = c(
Site_number="Site",
M.C="M_to_C",
BIX="BIX",
HIX="HIX",
FI="FI",
Ultra.Violet.Humic.like="CobleA",
Marine.Humic.like="CobleM",
Visible.Humic.like="CobleC",
Tryptophan.like="CobleT",
Tyrosine.like="CobleB",
Phenylalanine.like="Fpeak",
Lignin.like = "Lignin"
)



for(i in 1:length(name_key)){
 
  setnames(fixed_mo21_fdom,
           old = names(name_key)[i],
           new = unname(name_key)[i],
           skip_absent = T)
   
}


# These are the indices we want for the model
keep_names = c(
  "Site",
  "Year",
  "CobleA",
  "CobleB",
  "CobleC",
  "CobleM",
  "CobleT",
  "Lignin",
  "M_to_C",
  "HIX",
  "BIX",
  "FI"
)



# add a year column
fixed_mo21_fdom[, Year := rep(2021, nrow(fixed_mo21_fdom))]
fixed_mo22_fdom[, Year := rep(2022, nrow(fixed_mo22_fdom))]

# keep only our selected columns
fixed_mo21_fdom = fixed_mo21_fdom[, .SD, .SDcols = keep_names]
fixed_mo22_fdom = fixed_mo22_fdom[, .SD, .SDcols = keep_names]


# This is our final clean fdom sheet!
fixed_all_fdom = rbind(fixed_mo21_fdom,fixed_mo22_fdom)


# This is our final merged dataset
fdom_merged_export = merge(DataTurbWater, fixed_all_fdom, by = c("Site","Year"))

fwrite(fdom_merged_export, "data/clean/DatTurbWaterfDOM.csv")

```







