---
title: "ATI fDOM explore"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    smooth_scroll: true 
    fig_width: 10
    code_folding: hide  
date: '2022-12-05'
---
# Overviewi

This rmarkdown is a first pass looking at data from all of the ATI 2022 fDOM (as processed by Sean's Rscript translation of the original matlab script).

These samples were run by Skye Inn on a Horiba Aqualog and were split over several runs.

A couple things to consider:

- We suspect contaminant fluorescence in the region of PAHs/CobleT because of recent experiments with JP5 fuel.  

- Some samples were re-run across runs.
    - Check if these look the same across runs.
    - Remove the rerun sample if nothing looks amiss
    
- Some milliq blanks were run as a form of extra QC.
    - Check these for unusual fluoresecence.
    - Remove from data set containing only real samples.


  
```{r env-setup}
source("R/src/eem_plot_functions.R")
source("R/src/process_aqualog_functions.R")
library(GGally)
library(rnaturalearth)

# set ggplot params
theme_set(theme_minimal())
options(ggplot2.discrete.colour = c(
  '#636EFA',
  '#EF553B',
  '#00CC96',
  '#AB63FA',
  '#FFA15A',
  '#19D3F3',
  '#FF6692',
  '#B6E880',
  '#FF97FF',
  '#FECB52')
  )

# identify indices
# select HIX, A, T, B for highlight

all_indices = c(
  'CobleA',
  'CobleB',
  'CobleC',
  'CobleM',
  'CobleT',
  'Fpeak',
  'Stedmon_D',
  'Optical_Brighteners',
  'dieselBandII',
  'Petroleum',
  'Lignin',
  'BIX',
  'HIX',
  'FI',
  'M_to_C',
  'JP5_empirical'
)

ord_indices =c(
  'CobleA',
  'CobleB',
  'CobleT')

select_indices = c(
  'CobleA',
  'CobleB',
  'CobleT',
  'HIX'
)


```


```{r read-in}
all_runs = readRDS("data/compiled/all_runs.rds")

sam_sheets = fread("data/compiled/all_sample_sheets.csv", drop = "run_name")
index_data = fread("data/compiled/all_sample_indices.csv")
meta = fread("data/sample_data/islandwide_sites_allMetadata.csv")
nutrients = fread("data/sample_data/Lagoon_nutrients_April_2022.csv")
turb = fread("data/sample_data/Turb_CHN_compiled_April2022.csv")

setnames(turb, "Site_Number","Site")



proj = c(
    cabral = "SGD",
    varari = "SGD",
    RRCoral_1 = "RRCoral",
    RRCoral_2 = "RRCoral",
    spiffy = "Spiffy",
    spiffy_lang = "SpiffyLang",
    diel_run1 = "diel",
    diel_run2 = "diel",
    aroundtheisland_run1 = "ati",
    aroundtheisland_run2 = "ati",
    aroundtheisland_run3 = "ati",
    aroundtheisland_run4 = "ati",
    aroundtheisland_run5 = "ati",
    aroundtheisland_run6 = "ati",
    aroundtheisland_run7 = "ati",
    aroundtheisland_run8 = "ati"
)

# subset to ATI
index_data[ , project := proj[run_name]]

index_data = index_data[project == "ati"]


# merge with relevant data
all_dat = merge(index_data,
                sam_sheets,
                all.x = T,
                all.y = F,
                by = "UniqueID")


all_dat[ , Site:= gsub("_.*","",UniqueID)]


# pull out ancillary samples

opu = all_dat[Site == "opu"]
lter = all_dat[Site == "lter"]
milliq = all_dat[Site=="milliq"]
dupd = all_dat[duplicated(all_dat$Site)]


dat = all_dat[!(Site %in% c("opu","lter","milliq"))] 

# make long data for ggplot
long_dat = melt.data.table(dat,
                            id.vars = c("UniqueID", "Site"),
                            measure.vars = all_indices,
                            variable.name = "index_name",
                            value.name = "index_value")





```
  


## Check re-run samples

Some samples were re-run to check for consistency (i.e. contaminant signal, run effects). I want to check that the re-run samples look similar to the original samples. 

```{r check-samples}
# pull out duplicates
dupd_dat = dat[Site %in% dupd$Site]
long_dupd_dat = long_dat[ Site %in% dupd$Site]

# check counts
dupd_dat[ , .("N samples" = .N,
              "Runs" = paste(run_name, collapse = ",")), by = "Site"]

# plot indices for each pair of duplicate samples

p_check_list = list()

for( i in dupd$Site){
 
   p_check_list[[i]] = ggplot(long_dupd_dat[Site == i],
                              aes(x = index_name,
                                  y = index_value,
                                  color = UniqueID))+
                        geom_point(size = 4, alpha = 0.5)+
                        theme(axis.text.x = element_text(angle = 90))

  print(p_check_list[[i]])
   
   
   ggsave(plot = p_check_list[[i]],
          filename = paste0("plots/exploration/ATI/Site_",i,"_rerun_check.pdf"),
          width = 10, height = 8)
}




```


## Check Miliq blanks

Skye ran the ATI samples. Whenever she saw a blank that she thought 'looked weird' she would run some milliq samples to see if the contamination stuck around. I didn't really encourage this, since it was sporadic contamination anyways, but I figured we should take a look at these data since we have them. 

```{r}
long_milliq = melt.data.table(milliq,
                            id.vars = c("UniqueID", "Site"),
                            measure.vars = all_indices,
                            variable.name = "index_name",
                            value.name = "index_value")

milliq_samples = unique(long_milliq$UniqueID)

 
   p_milliq = ggplot(long_milliq,
                              aes(x = index_name,
                                  y = index_value))+
                        geom_point(size = 4, alpha = 0.5)+
                        theme(axis.text.x = element_text(angle = 90))

  
   ggsave(plot = p_milliq,
          filename = paste0("plots/exploration/ATI/Site_",i,"_milliq_check.pdf"),
          width = 10, height = 8)
p_milliq

```

## Remove extraneous reruns and blanks

Now that we've checked for consistency between re-run samples and the opportunistically run milliq blanks, we can remove them from the data set and move onto analyses on the real data.


```{r}

clean_dat = merge(meta, 
                  dat[!(UniqueID %in% dupd$UniqueID),],
                  by = "Site",
                  all.y = T)

clean_dat = merge(clean_dat, 
                  nutrients,
                  by = "Site",
                  all.X = T)

clean_dat = merge(clean_dat,
                  turb,
                  by = "Site",
                  all.X = T)

fwrite(clean_dat, "data/clean/ati_fdom_clean.csv")


clean_long_dat = melt.data.table(clean_dat,
                            measure.vars = all_indices,
                            variable.name = "index_name",
                            value.name = "index_value")

fwrite(clean_long_dat, "data/clean/ati_fdom_long.csv")

```



# Site Data

Check for trends by categorical variables:

1. Habitat
2. shore direction
3. Lagoon name

This is just a rough first pass with ugly boxplots.

Check for trends with Nutrients
1. Phosphate
2. Silicate
3. N+N
4. Ammonia

## Categoricals


```{r}
# check meta categories

meta[ , .N, by = .(Habitat)][order(N)]

meta[ , .N, by = .(Island_shore)][order(N)]

meta[ , .N, by = .(Lagoon)][order(N)]


# check nutrients
nutrient_vars = c("Phosphate",
                  "Silicate",
                  "Nitrite_plus_Nitrate",
                  "Ammonia")

ggpairs(nutrients, 3:6)



```


```{r}
# habitat

#pdf("plots/exploration/Habitat_index_plots.pdf", width = 10, height = 8)

for(ind in unique(clean_long_dat$index_name)){
  
p = ggplot(clean_long_dat[index_name == ind], aes(x = index_name,
                                         y = log(index_value),
                                         color = Habitat))+
  geom_boxplot()+
    labs(title = paste0("log ",ind))

print(p)
  
  
}
#dev.off()



```


```{r}
# shore

#pdf("plots/exploration/Island_shore_index_plots.pdf", width = 10, height = 8)

for(ind in unique(clean_long_dat$index_name)){
  
p = ggplot(clean_long_dat[index_name == ind], aes(x = index_name,
                                         y = log(index_value),
                                         color = Island_shore))+
  geom_boxplot()+
    labs(title = paste0("log ",ind))

print(p)
  
  
}
#dev.off()

```

  
```{r}
#pdf("plots/exploration/Lagoon_index_plots.pdf", width = 10, height = 8)

for(ind in unique(clean_long_dat$index_name)){
  
p = ggplot(clean_long_dat[index_name == ind], aes(x = index_name,
                                         y = log(index_value),
                                         color = Lagoon))+
  geom_boxplot()+
    labs(title = paste0("log ",ind))

print(p)
  
  
}
#dev.off()

```


## Nutrients

```{r}

```



## Maps  


## version 1
Working on a basi ggplot version of maps, but the projection seems to be off for some reason. Everything is in WGS84, so not sure what that's about.
  
```{r}
library(raster)
library(sf)
library(sp)

# remove sites with no lat/lon
map_meta = meta[!is.na(Latitude)]

map_long_dat = clean_long_dat[!is.na(Latitude)]

# make spacial points data frame
map_sp = SpatialPointsDataFrame(data = map_long_dat,
                       coords = list(map_long_dat$Latitude,
                                  map_long_dat$Longitude))


# Check geographic range of sampling points
limits = c(
  min(map_meta$Longitude),
  min(map_meta$Latitude), 
  max(map_meta$Longitude),
  max(map_meta$Latitude) 
)

# define a bounding box with a small cushion around the minimum and maximum
bbox = list(
  xmin = limits[1] - 0.03,
  ymin = limits[2] - 0.04,
  xmax = limits[3] + 0.03,
  ymax = limits[4] + 0.04
)

# load shape file
coasts = st_read("data/gis/ne_10m_coastline.shp")


mo_coast = st_crop(coasts,
        xmin = bbox$xmin,
        xmax = bbox$xmax,
        ymin = bbox$ymin,
        ymax = bbox$ymax)

crs(map_sp)


# test
ggplot()+
    geom_point(data = as.data.frame(map_sp),
       aes(x = Longitude, y = Latitude, color = index_value))+
      geom_sf(data = mo_coast)


```


## version 2
  This is another attempt at maps using the OpenStreet package. Seems to handle the point projection correctly, so I'm happy!
  
  All the indices are log transformed to better show spread at low values. There are a couple outlier samples with very high values for certain indices that wash out the colors otherwise. 
  
  Currently, these maps just show colored points. Eventually, we could try some interpolation or something. 
  
  
```{r}
library(OpenStreetMap)
library(viridisLite)

sa_map <- openmap(c(bbox$ymax, bbox$xmin),
                  c(bbox$ymin, bbox$xmax),
                  type = "stamen-terrain",
                  mergeTiles = TRUE)

sa_map2 <- openproj(sa_map)


for(an_index in unique(map_long_dat$index_name)){

sa_map2_plt <- OpenStreetMap::autoplot.OpenStreetMap(sa_map2)+
  geom_point(data = map_long_dat[index_name == an_index],
             aes(x = Longitude,
                 y = Latitude,
                 color = log(index_value)))+
  labs(title = paste("Log",an_index))+
  scale_color_viridis_c()

print(sa_map2_plt)

ggsave(filename = paste0("plots/exploration/ATI/map_",an_index,".pdf"), plot = sa_map2_plt)
  
}

```
  
  