---
title: "ATI fDOM explore"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    smooth_scroll: true 
    fig_width: 10
    code_folding: hide  
date: '2022-12-05'
---
# Overview

This rmarkdown is a first pass looking at data from all of the ATI 2022 fDOM (as processed by Sean's Rscript translation of the original matlab script).

These samples were run by Skye Inn on a Horiba Aqualog and were split over several runs.

A couple things to consider:

- We suspect contaminant fluorescence in the region of PAHs/CobleT because of recent experiments with JP5 fuel.  

- Some samples were re-run across runs.
    - Check if these look the same across runs.
    - Remove the rerun sample if nothing looks amiss
    
- Some milliq blanks were run as a form of extra QC.
    - Check these for unusual fluoresecence.
    - Remove from data set containing only real samples.


  
```{r env-setup}
source("R/src/eem_plot_functions.R")
source("R/src/process_aqualog_functions.R")
library(GGally)
library(oce)

# set ggplot params
theme_set(theme_minimal())
options(ggplot2.discrete.colour = c(
  '#636EFA',
  '#EF553B',
  '#00CC96',
  '#AB63FA',
  '#FFA15A',
  '#19D3F3',
  '#FF6692',
  '#B6E880',
  '#FF97FF',
  '#FECB52')
  )

# identify indices

all_indices = c(
  'CobleA',
  'CobleB',
  'CobleC',
  'CobleM',
  'CobleT',
  'Fpeak',
  'Stedmon_D',
  'Optical_Brighteners',
  'dieselBandII',
  'Petroleum',
  'Lignin',
  'BIX',
  'HIX',
  'FI',
  'M_to_C',
  'JP5_empirical'
)

ord_indices =c(
  'CobleA',
  'CobleB',
  'CobleT',
  'Lignin'
)

```


```{r read-in}
all_runs = readRDS("data/compiled/all_runs.rds")

sam_sheets = fread("data/compiled/all_sample_sheets.csv", drop = "run_name")
index_data = fread("data/compiled/all_sample_indices.csv")
site_data = fread("data/sample_data/islandwide_sites_allMetadata.csv")



proj = c(
    cabral = "SGD",
    varari = "SGD",
    RRCoral_1 = "RRCoral",
    RRCoral_2 = "RRCoral",
    spiffy = "Spiffy",
    spiffy_lang = "SpiffyLang",
    diel_run1 = "diel",
    diel_run2 = "diel",
    aroundtheisland_run1 = "ati",
    aroundtheisland_run2 = "ati",
    aroundtheisland_run3 = "ati",
    aroundtheisland_run4 = "ati",
    aroundtheisland_run5 = "ati",
    aroundtheisland_run6 = "ati",
    aroundtheisland_run7 = "ati",
    aroundtheisland_run8 = "ati"
)

# subset to ATI
index_data[ , project := proj[run_name]]

index_data = index_data[project == "ati"]


# merge with relevant data
all_dat = merge(index_data,
                sam_sheets,
                all.x = T,
                all.y = F,
                by = "UniqueID")
all_dat[ , Site:= gsub("_.*","",UniqueID)]


# pull out ancillary samples

opu = all_dat[Site == "opu"]
lter = all_dat[Site == "lter"]
milliq = all_dat[Site=="milliq"]

dat = all_dat[!(Site %in% c("opu","lter","milliq"))] 

# make long data for ggplot
long_dat = melt.data.table(dat,
                            id.vars = c("UniqueID", "Site"),
                            measure.vars = all_indices,
                            variable.name = "index_name",
                            value.name = "index_value")


# pull out duplicates

dupd_samples = dat$Site[duplicated(dat$Site)]

dupd_dat = dat[Site %in% dupd_samples]
long_dupd_dat = long_dat[ Site %in% dupd_samples]


```
  


## Check re-run samples

Some samples were re-run to check for consistency (i.e. contaminant signal, run effects).

```{r check-samples}
# check counts
dupd_dat[ , .("N samples" = .N,
              "Runs" = paste(run_name, collapse = ",")), by = "Site"]

# plot indices for each pair

p_check_list = list()

for( i in dupd_samples){
 
   p_check_list[[i]] = ggplot(long_dupd_dat[Site == i],
                              aes(x = index_name,
                                  y = index_value,
                                  color = UniqueID))+
                        geom_point(size = 4, alpha = 0.5)+
                        theme(axis.text.x = element_text(angle = 90))

  
   ggsave(plot = p_check_list[[i]],
          filename = paste0("plots/exploration/ATI/Site_",i,"_rerun_check.pdf"),
          width = 10, height = 8)
}




```


## Check Miliq blanks

```{r}
long_milliq = long_dat[Site == "milliq"]


```

## Remove extraneous reruns and blanks

```{r}



```



# Site Data

Check for trends by:

1. Habitat
2. shore direction
3. Lagoon name
4. Lat/Lon


```{r}

meta = fread("data/sample_data/islandwide_sites_allMetadata.csv")

meta[ , .N, by = .(Habitat)]

meta[ , .N, by = .(Island_shore)]

meta[ , .N, by = .(Lagoon)]


```

  
  
  