---
title: "Mo22 fDOM explore"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    smooth_scroll: true 
    fig_width: 10
    code_folding: hide  
date: '2022-12-05'
---
# Overview

This rmarkdown compiles all of the Moorea 2022 fDOM data (as processed by Sean's Rscript translation of the original matlab script).


## Projects

Below is a list of projects and runs included in this analysis.
All fDOM was run by Sean, except for Around the Island, which was run by Skye Inn.

* RRCoral (Catherine)
  * RRC_1
  * RRC_2

* Spiffy (Ralph/Irina)
  * Spiffy
  
* Spiffy Lang (Ralph/Irina)
  *SL

* Diel (Wegeley/Nelson/Haas)
  *Diel_1
  *Diel_2
  
* Around the Island (Burkepile Lab)
  * ATI_1
  * ATI_2
  * ATI_3
  * ATI_4
  * ATI_5
  * ATI_6
  * ATI_7
  * ATI_8
  

Other runs that were part of Moorea 2022 but analyzed separately:

* SGD
  Cabral + Varari
  Processed with matlab script to maintain consistency with prior data.
  
* TURNT
  Processed with R script.
  
  

  
  
```{r}
source("R/src/eem_plot_functions.R")
source("R/src/process_aqualog_functions.R")
library(GGally)

theme_set(theme_minimal())
options(ggplot2.discrete.colour= c('#636EFA', '#EF553B', '#00CC96', '#AB63FA', '#FFA15A', '#19D3F3', '#FF6692', '#B6E880', '#FF97FF', '#FECB52'))

all_runs = readRDS("data/compiled/all_runs.rds")

sam_sheets = fread("data/compiled/all_sample_sheets.csv")
index_data = fread("data/compiled/all_sample_indices.csv")


proj = c(
    cabral = "SGD",
    varari = "SGD",
    RRCoral_1 = "RRCoral",
    RRCoral_2 = "RRCoral",
    spiffy = "Spiffy",
    spiffy_lang = "SpiffyLang",
    diel_run1 = "diel",
    diel_run2 = "diel",
    aroundtheisland_run1 = "ati",
    aroundtheisland_run2 = "ati",
    aroundtheisland_run3 = "ati",
    aroundtheisland_run4 = "ati",
    aroundtheisland_run5 = "ati",
    aroundtheisland_run6 = "ati",
    aroundtheisland_run7 = "ati",
    aroundtheisland_run8 = "ati"
)

index_data[ , project := proj[run_name]]


```
  
  
  
  
  
  
  
```{r}
# identify indices for plots and ordination

indices = c("CobleA",
            "CobleB",
            "Lignin",
            "CobleT")

check_index = "HIX"


# standardized ordination function
ordinate_dt = function(dt,
                       id_col = "UniqueID",
                       cols = indices){
  require(vegan)
  set.seed("2022")
  
  ord_dat = as.matrix(dt[ , ..cols])
  row.names(ord_dat) = dt[[id_col]]
  
  # drop samples that won't produce a valid distance
  row_check = apply(ord_dat,
                    1,
                    function(x) any(is.na(x) |
                                    is.infinite(x)|
                                    sum(x) == 0))
  
  drop_samples = names(row_check)[row_check]
  
  warning(
    paste("Dropping samples with NA, Inf, or all zero values: ",
                paste("\n", drop_samples, collapse = "\r"))
    )
  
  ord_dat = ord_dat[!(row.names(ord_dat) %in% drop_samples),]
  
    
  # ordinate
  raw_ord = metaMDS(ord_dat, distance = "bray",  autotransform = F)
  
  ord_points = as.data.table(raw_ord$points, keep.rownames = T)
  ord_points[ , d_from_0 := sqrt((MDS1)^2+(MDS2)^2)]
  setnames(ord_points, "rn", id_col)
  
  ord_points = merge(ord_points, dt, all.x=T, by = id_col)
  
  return(
    list( points = ord_points,
          stress = raw_ord$stress)
    )
  
}
```


```{r}
# main qc function
qc_fdom = function(dat,
                   id_col = "UniqueID",
                   color_col = "run_name",
                   indices,
                   check_index = NULL){
  
  # organize data
  keep_cols = c(id_col, color_col, indices, check_index)
  
  dat = dat[ , ..keep_cols]
  
  long_dat = melt.data.table(dat,
                             id.vars = c(id_col,color_col),
                             measure.vars = c(indices, check_index),
                             variable.name = "index_name",
                             value.name = "index_value")
  
  long_dat$log_index_value = log(long_dat$index_value)
  
  # summarize data
  p_counts = ggplot(dat,
                    aes_string(x = color_col,
                               color = color_col))+
              geom_bar(fill = "white")+
              labs(title = "N samples per run")+
              theme(axis.text.x = element_text(angle = 90))
  
  # pair plot
  p_hist = ggpairs(dat, columns = indices)
  
  
  # comparative violin plots
  p_box = ggplot(long_dat, aes_string(x = "index_name",
                               y = "log_index_value",
                               color = color_col))+
    #geom_jitter(alpha = 0.2)+
    geom_boxplot(alpha = 0.7)
  
  # check index violin plot
  if(!is.null(check_index)){
  p_check = ggplot(long_dat[index_name %in% check_index,],
                   aes_string(y = "index_value",
                              x = color_col,
                              color = color_col))+
              geom_boxplot(outlier.shape = NA)+
              geom_jitter(alpha = 0.5, pch = 19)+
              labs(y = check_index, title = paste("Check",
                                                  check_index,
                                                  "Across Runs"))+
              theme(axis.text.x = element_text(angle = 90))
  }

  # ordination plot
  ord_points = ordinate_dt(dat,
                           id_col = id_col,
                           cols = indices)
 
  outliers = tail(ord_points$points[order(d_from_0)])
  y_range = range(ord_points$points$MDS1)
  y_offset = (y_range[2]-y_range[1])/50
 
  p_ord = ggplot(ord_points$points, aes(x = MDS1, y = MDS2, color = run_name))+
    geom_point(alpha = 0.7)+
    geom_point(data = outliers,
               aes(x = MDS1,
                   y = MDS2),
               pch = 4,
               alpha = 0.5)+
    geom_text(data = outliers,
               aes(x = MDS1,
                   y = MDS2,
                   label = UniqueID),
              nudge_y = y_offset
               )
  out = list(
    p_counts = p_counts,
    p_hist = p_hist,
    p_box = p_box,
    p_check = p_check,
    p_ord = p_ord,
    ord_points = ord_points
  )
  
  
  return(out) 
  
}


```
  
  
```{r}
#QC all mo22 fdom runs

all_runs = qc_fdom(dat = index_data, 
        indices = indices,
        id_col = "UniqueID",
        check_index = check_index)


pdf("plots/exploration/all_runs_qc_plots.pdf",
    width = 10,
    height = 8)
  print(all_runs)
dev.off()

# for each project

runs = unique(index_data$run_name)

proj_qc = list()

for(i in unique(proj)){
 proj_dat = index_data[project %in% i]
 proj_qc[[i]] = qc_fdom(dat = proj_dat,
                        indices = indices,
                        id_col = "UniqueID",
                        check_index = check_index)
}

for(i in 1:length(proj_qc)){
  pdf(paste0("plots/exploration/",names(proj_qc)[i],"_qc_plots.pdf"),
      width = 10, height = 8)
  print(proj_qc[[i]])
  dev.off()
   
}



```
  
  
  
  
  